---
title: "LP OPTIMIZATION WITH R"
output: html_document
---

```{r}
library(lpSolve)
library(lpSolveAPI)

```
###################################TABLE CHAIR PRODUCTION#################################

•A manufacturer produces two pieces of furniture: tables and chairs

–  Tables sell for $1600 per piece and chairs sell for $1000 per pieces

•Production of furniture requires two types of raw-material (lego pieces): small and large pieces

–  There are 6 large and 8 small pieces available

•Determine optimal production plan that maximizes total revenue

Excel solution 
Decision Variable	Table-x1	Chair-x2			
	2	2			
Obj. Coef.	1600	1000	OBJ=	5200	
Consts	2	1	6	<=	6
	2	2	8	<=	8

#SOLUTION 1 FOR CHAIR and TABLE QUESTION
```{r}
obj.fun <- c(1600,1000)
constr <- matrix (c(2,1,2,2) , ncol = 2, byrow =TRUE)
constr.dir<- c("<=", "<=")
rhs <- c(6,8)


prod.sol <- lp("max", obj.fun , constr , constr.dir , rhs ,compute.sens = TRUE )


prod.sol$ objval #objective function value
prod.sol$ solution #decision variables values
prod.sol$ duals #includes duals of constraints and reduce costs of variables, gives shodow price then rduce cost.



```
#SOLUTION 2 FOR CHAIR and TABLE QUESTION
```{r}

library(lpSolveAPI)

lpmodel <- make.lp(0,2) # there are 2 decision variables
lp.control(lpmodel, sense="max")
set.objfn(lpmodel, c(1600,1200))
add.constraint(lpmodel, c(2,1), "<=", 6)
add.constraint(lpmodel, c(2,2), "<=", 8)


solve(lpmodel)
get.objective(lpmodel)
get.variables(lpmodel)
get.dual.solution(lpmodel)

lpmodel

```

###############################Product 1 and Product 2#####################################

A small business sells two products, named Product 1 and Product 2. Each tonne of Product 1 consumes 30 working hours, and each tonne of Product 2 consumes 20 working hours. The business has a maximum of 2,700 working hours for the period considered. As for machine hours, each tonne of Products 1 and 2 consumes 5 and 10 machine hours, respectively. There are 850 machine hours available.

Each tonne of Product 1 yields 20 Me of profit, while Product 2 yields 60 Me for each tonne sold. For technical reasons, the firm must pro-duce a minimum of 95 tonnes in total between both products. We need to know how many tonnes of Product 1 and 2 must be produced to maximize total profit.

**Below is excel solution
Decision Variable	product1	product2			
	20	75			
Obj. Coef.	20	60	OBJ=	4900	
Consts	30	20	2100	<=	2700
	5	10	850	<=	850
	1	1	95	>=	95



#SOLUTION1 FOR Product Problem
```{r eval=FALSE}
install.packages("lpSolveAPI")
library(lpSolveAPI)

lpmodel <- make.lp(0,2)
lp.control(lpmodel, sense="max")
set.objfn(lpmodel, c(20,60))
add.constraint(lpmodel, c(30,20), "<=", 2700)
add.constraint(lpmodel, c(5,10), "<=", 850)
add.constraint(lpmodel, c(1,1), ">=", 95)


solve(lpmodel)
get.objective(lpmodel) #4900 optimized profit
get.variables(lpmodel) # solution  20    60  
get.dual.solution(lpmodel) # dual(shadow price) 1   0   8 -20   reduced cost: 0   0

lpmodel

```
#SOLUTION2 FOR Product Problem
```{r eval=FALSE}
#the same problem LPSOLve package solution,#example 2 solution with lpSolve
library(lpSolve)

#defining parameters
obj.fun <- c(20, 60)
constr <- matrix(c(30, 20, 5, 10, 1, 1), ncol = 2, byrow=TRUE)
constr.dir <- c("<=", "<=", ">=") 
rhs <- c(2700, 850, 95)

#solving model
prod.sol <- lp("max", obj.fun, constr, constr.dir, rhs, compute.sens = TRUE)
prod.sol
#accessing to R output
prod.sol$objval #objective function value 
prod.sol$solution #decision variables values 
prod.sol$duals #includes duals of constraints and reduced costs of variables

#sensitivity analysis results
prod.sol$duals.from
prod.sol$duals.to
prod.sol$sens.coef.from
prod.sol$sens.coef.to

```

###############################DESTINATION PROBLEM#####################################

Let’s consider a transportation problem of two origins a and b, and three destinations 1, 2 and 3. In Table 2.1 are presented the cost cij of trans-porting one unit from the origin i to destination j, and the maximal capacity of the origins and the required demand in the destinations. We need to know how we must cover the demand of the destinations at a minimal cost.



1	2	3	capacity
				
a	8	6	3	70
b	2	4	9	40
				
demand	40	35	25	
				


This situation can be modeled with a LP with the following elements:

•Decision variables of the form xij, representing units transported from origin i to destination j

•An objective function with cost coefficients equal to cij

•Two sets of constraints: a less or equal set of constraints for each origin, limiting the units to be transported, and a greater of equal set of constraints representing that the demand of each destina-tion must be covered.

The resulting LP is:

MIN z = 8xa1 + 6xa2 + 3xa3 + 2xb1 + 4xb2 + 9xb3 

s.a. ca) xa1 + xa2 + xa3 70

cb) xb1 + xb2 + xb3	40

d1) xa1 + xb1	40

d2) xa2 + xb2	35

d3) xa3 + xb3	25

xij	0

EXCEL SOLVER SOLUTION is also below

Decision Variable	a1	a2	a3	b1	b2	b3		
	                 0	35	25	40	0	  0		
Obj. Coef.	8	6	3	2	4	9	OBJ=	365
Consts	1	1	1	60	<=	70		
	1	1	1	40	<=	40		
	1	1		40	>=	40		
	1	1		35	>=	35		
	1	1		25	>=	25		

#SOLUTION 1
```{r eval=FALSE}
#defining parameters
#origins run i in 1:m #destinations run j in 1:n 
library(lpSolve)
obj.fun <- c(8, 6, 3, 2, 4, 9)
m <- 2
n <- 3
constr <- matrix(0, n+m, n*m)
for(i in 1:m){ 
  for(j in 1:n){ 
    constr[i, n*(i-1) + j] <- 1 
    constr[m+j, n*(i-1) + j] <- 1 
  } 
}

constr.dir <- c(rep("<=", m), rep(">=", n)) 
rhs <- c(70, 40, 40, 35, 25) 

#solving LP model 
prod.trans <- lp("min", obj.fun, constr, constr.dir, rhs, compute.sens = TRUE) 

#LP solution 
prod.trans$objval
sol <- matrix(prod.trans$solution , m, n, byrow=TRUE)
prod.trans$duals

#sensitivity analysis of LP 
prod.trans$duals.from 
prod.trans$duals.to 
prod.trans$sens.coef.from 
prod.trans$sens.coef.to 
```

#SOLUTION 2

```{r eval=FALSE}
install.packages("lpSolveAPI")
library(lpSolveAPI)

lprec <- make.lp(0, 6) #there are 6 decision variables
lp.control(lprec, sense="min")
set.objfn(lprec, c(8, 6, 3, 2, 4, 9)) #objective function
add.constraint(lprec, c(1, 1, 1,0,0,0), "<=", 70)
add.constraint(lprec, c(0,0,0,1,1,1), "<=", 40)
add.constraint(lprec, c(1,0,0, 1,0,0), ">=", 40)
add.constraint(lprec, c(0,1,0, 0,1,0), ">=", 35)
add.constraint(lprec, c(0,0,1, 0,0,1), ">=", 25)
lprec
solve(lprec)
#maximum profit
get.objective(lprec)
#Get the solution
get.variables(lprec)
matrix_sol <- matrix(get.variables(lprec), 2, 3, byrow=TRUE)
```

###################################CARGO EXAMPLE#################################


The example case;

A trading company is looking for a way to maximize profit per transportation of their goods. The company has a train available with 3 wagons. When stocking the wagons they can choose between 4 types of cargo, each with its own specifications. How much of each cargo type should be loaded on which wagon in order to maximize profit?


The following constraints have to be taken in consideration;

Weight capacity per train wagon
Volume capacity per train wagon
Limited availability per cargo type
Let’s assume we have the following information at hand:

TRAIN WAGON	WEIGHT CAPACITY (TONNE)	SPACE CAPACITY (M2)
w1	10	5000
w2	8	4000
w3	12	8000
CARGO TYPE	AVAILABLE (TONNE)	VOLUME (M2)	PROFIT (PER TONNE)
c1	18	400	2000
c2	10	300	2500
c3	5	200	5000
c4	20	500	3500

#SOLUTION 1 FOR CARGO QUESTION
```{r eval=FALSE}
library(lpSolveAPI)
install.packages("reshape")
install.packages("gridExtra")
 
#used for result visualization
library(ggplot2)
library(reshape)
library(gridExtra)
 
#define the datasets
  
train<-data.frame(wagon=c('w1','w2','w3'), weightcapacity=c(10,8,12), spacecapacity=c(5000,4000,8000))
 
cargo<-data.frame(type=c('c1','c2','c3','c4'), available=c(18,10,5,20), volume=c(400,300,200,500),profit=c(2000,2500,5000,3500))

#create an LP model with 10 constraints and 12 decision variables
 
lpmodel<-make.lp(2*NROW(train)+NROW(cargo),12)
 
column<-0
row<-0
 
#build the model column per column
for(wg in train$wagon){
    row<-row+1
    for(type in seq(1,NROW(cargo$type))){
    column<-column+1
     
    #this takes the arguments 'column','values' & 'indices' (as in where these values should be placed in the column)
    set.column(lpmodel,column,c(1, cargo[type,'volume'],1), indices=c(row,NROW(train)+row, NROW(train)*2+type))
    }}
 
#set rhs weight constraints
set.constr.value(lpmodel, rhs=train$weightcapacity, constraints=seq(1,NROW(train)))
 
#set rhs volume constraints
set.constr.value(lpmodel, rhs=train$spacecapacity, constraints=seq(NROW(train)+1,NROW(train)*2))
 
 
#set rhs volume constraints
set.constr.value(lpmodel, rhs=cargo$available, constraints=seq(NROW(train)*2+1,NROW(train)*2+NROW(cargo)))
 
#set objective coefficients
set.objfn(lpmodel, rep(cargo$profit,NROW(train)))
 
#set objective direction
lp.control(lpmodel,sense='max')
 
#I in order to be able to visually check the model, I find it useful to write the model to a text file
write.lp(lpmodel,'model.lp',type='lp')
#solve the model, if this return 0 an optimal solution is found
solve(lpmodel)
 
#this return the proposed solution
get.objective(lpmodel)
```

#SOLUTION 2 FOR CARGO QUESTION
```{r eval=FALSE}
# Load the lpSolveAPI library
library(lpSolveAPI)

DataForModel <- read.table("CargoData.csv", sep=",", header=TRUE)

NoOfConstraints <- ncol(DataForModel)-2
NoOfVariables <-nrow(DataForModel)-1

# We create a new empty model 
lpModel <-make.lp(NoOfConstraints,NoOfVariables)

#Name the model variables
VariableNames <- t(DataForModel[1:NoOfVariables,1])
#Name the constraints
ConstraintNames <- colnames(DataForModel[3:(NoOfConstraints+2)])

# minimize the cost 
lp.control(lpModel, sense="max")

for(i in 1:NoOfVariables){ 
  set.column(lpModel, i, DataForModel[i,3:(NoOfConstraints+2)])
} 
# Each constraint provides the lower bound of rhs
set.constr.type(lpModel,rep("<=", NoOfConstraints))


# Set the RHS values
set.rhs(lpModel, DataForModel[(NoOfVariables+1),3:(NoOfConstraints+2)])


dimnames(lpModel) <- list(ConstraintNames, VariableNames)

#Objective
obj.profit <- DataForModel[1:NoOfVariables,2]
set.objfn(lpModel, obj.profit)
lpModel
# In order to be able to visually check the model, you can write the model to a text file
write.lp(lpModel,'modelCargo.lp',type='lp')

# Solving the model
solve(lpModel) 
get.objective(lpModel)
get.variables(lpModel)

```


#SOLUTION 3 FOR CARGO QUESTION
```{r eval=FALSE}
lpmodel <- make.lp(0,12)
lp.control(lpmodel, sense="max")
set.objfn(lpmodel, c(2000,2000,2000,2500,2500,2500,5000,5000,5000,3500,3500,3500))

add.constraint(lpmodel, c(1,0,0,1,0,0,1,0,0,1,0,0), "<=", 10)
add.constraint(lpmodel, c(0,1,0,0,1,0,0,1,0,0,1,0), "<=", 8)
add.constraint(lpmodel, c(0,0,1,0,0,1,0,0,1,0,0,1), "<=", 12)

add.constraint(lpmodel, c(400,0,0,300,0,0,200,0,0,500,0,0), "<=", 5000)
add.constraint(lpmodel, c(0,400,0,0,300,0,0,200,0,0,500,0), "<=", 4000)
add.constraint(lpmodel, c(0,0,400,0,0,300,0,0,200,0,0,500), "<=", 8000)

add.constraint(lpmodel, c(1,1,1,0,0,0,0,0,0,0,0,0), "<=", 18)
add.constraint(lpmodel, c(0,0,0,1,1,1,0,0,0,0,0,0), "<=", 10)
add.constraint(lpmodel, c(0,0,0,0,0,0,1,1,1,0,0,0), "<=", 5)
add.constraint(lpmodel, c(0,0,0,0,0,0,0,0,0,1,1,1), "<=", 20)

solve(lpmodel)
get.objective(lpmodel)
get.variables(lpmodel)
get.dual.solution(lpmodel)

lpmodel


```
###################################POST OFFICE EMPOYEE EXAMPLE###########################


#SOLUTION 1 FOR POST OFFICE

```{r eval=FALSE}

```